How to Split Data to Speed Up MERT

### 2012-02-16 doc/latex/paper.tex
Minor update to documentation to add citation to Schwartz et al. (2011)

\begin{figure*}
\centering
\subfloat[Decoding times in \textbf{seconds} for standard decoder configured using a 5-gram language model.]{
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={160,150,140,130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Seconds),
ymin=0,
%legend style={
%	at={(0.5,-0.15)},
%	anchor=north,legend columns=-1
%},
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{ngram/ngram.naive-Maximum_times}
\input{ngram/ngram.histogram-Maximum_times}
\input{ngram/ngram.words-Maximum_times}
\input{ngram/ngram.time-Maximum_times}
%
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model in addition to a 5-gram language model.]{
\begin{tikzpicture}%[scale=2]
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Hours),
ymin=0,
ybar=0pt,
bar shift=0pt
]
%
\input{hhmm/hhmm.naive-Maximum_times}
\input{hhmm/hhmm.histogram-Maximum_times}
\input{hhmm/hhmm.words-Maximum_times}
\input{hhmm/hhmm.time-Maximum_times}
%
%\legend{Naive,Histogram,Words,Oracle}
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{Decoding times for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms ({\sc Naive-Split}, {\sc Histogram-Split}, {\sc Words-Split}, and {\sc Times-Split}).}
\label{fig:maxspeed}
\end{figure*}

\begin{figure*}
\centering
\subfloat[Decoding times in \textbf{seconds} for standard decoder configured using a 5-gram language model.]{
\input{ngram/ngram.max-min-table.tex}
} \\
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model in addition to a 5-gram language model.]{
\input{hhmm/hhmm.max-min-table.tex}
}
\caption{Decoding times for the fastest (min) and slowest (max) decoding jobs when a translation task is split into $n$ decoding jobs. \textit{Italics} indicate balanced task times. \textbf{Bold} indicates fastest max time at that split.}
\label{fig:minmaxspeed}
\end{figure*}

\begin{figure*}[!tb]
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Seconds} for standard decoder configured using a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
scaled y ticks=false,
ytick={2000,1800,1600,1400,1200,1000,800,600,400,200,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Seconds,
ymin=0,
legend pos=north west,
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{ngram/ngram.naive-slack_times}
\input{ngram/ngram.histogram-slack_times}
\input{ngram/ngram.words-slack_times}
\input{ngram/ngram.time-slack_times}
%
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Hours} for decoder configured using a syntactic language model in addition to a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
ytick={3000,2500,2000,1500,1000,750,500,250,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Hours,
ymin=0,
legend pos=north west,
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{hhmm/hhmm.naive-slack_times}
\input{hhmm/hhmm.histogram-slack_times}
\input{hhmm/hhmm.words-slack_times}
\input{hhmm/hhmm.time-slack_times}
%
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{Cumulative slack CPU time for $n$ processing cores when processing a parallel translation task split into $n$ jobs using various splitting algorithms. Slack CPU time is caused when some jobs finish before others. Zero slack time indicates conditions where all jobs complete simultaneously.}
\label{fig:slack}
\end{figure*}

@@

To observe the effects of splitting algorithms on decoding speed, we translated Urdu-English data using Moses in a parallel computing cluster, distributing work using the Sun Grid Engine. We ran two decoding setups: a standard configuration using a 5-gram language model, and a much slower configuration that also used an incremental syntactic language model. Using Algorithm \ref{alg:naive}, the runtimes of the slowest of $n$ translation jobs in each configuration is illustrated in Figure \ref{fig:maxspeed} for various values of $n$. Figure \ref{fig:minmaxspeed} shows that in all cases, there is a significant difference between the fastest and slowest translation job, leading to computational slack time (Figure \ref{fig:slack}).

>>>>

\begin{figure*}
\centering
\subfloat[Decoding times in \textbf{seconds} for standard decoder configured using a 5-gram language model.]{
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={160,150,140,130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Seconds),
ymin=0,
%legend style={
%	at={(0.5,-0.15)},
%	anchor=north,legend columns=-1
%},
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{ngram/ngram.naive-Maximum_times}
\input{ngram/ngram.histogram-Maximum_times}
\input{ngram/ngram.words-Maximum_times}
\input{ngram/ngram.time-Maximum_times}
%
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}%[scale=2]
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Hours),
ymin=0,
ybar=0pt,
bar shift=0pt
]
%
\input{hhmm/hhmm.naive-Maximum_times}
\input{hhmm/hhmm.histogram-Maximum_times}
\input{hhmm/hhmm.words-Maximum_times}
\input{hhmm/hhmm.time-Maximum_times}
%
%\legend{Naive,Histogram,Words,Oracle}
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{Decoding times for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms ({\sc Naive-Split}, {\sc Histogram-Split}, {\sc Words-Split}, and {\sc Times-Split}).}
\label{fig:maxspeed}
\end{figure*}

\begin{figure*}
\centering
\subfloat[Decoding times in \textbf{seconds} for standard decoder configured using a 5-gram language model.]{
\input{ngram/ngram.max-min-table.tex}
} \\
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) addition to a 5-gram language model.]{
\input{hhmm/hhmm.max-min-table.tex}
}
\caption{Decoding times for the fastest (min) and slowest (max) decoding jobs when a translation task is split into $n$ decoding jobs. \textit{Italics} indicate balanced task times. \textbf{Bold} indicates fastest max time at that split.}
\label{fig:minmaxspeed}
\end{figure*}

\begin{figure*}[!tb]
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Seconds} for standard decoder configured using a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
scaled y ticks=false,
ytick={2000,1800,1600,1400,1200,1000,800,600,400,200,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Seconds,
ymin=0,
legend pos=north west,
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{ngram/ngram.naive-slack_times}
\input{ngram/ngram.histogram-slack_times}
\input{ngram/ngram.words-slack_times}
\input{ngram/ngram.time-slack_times}
%
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%
\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\subfloat[Slack \textbf{CPU-Hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={2,4,8,16,32,64,128,256,512},
ytick={3000,2500,2000,1500,1000,750,500,250,100},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Slack CPU-Hours,
ymin=0,
legend pos=north west,
ybar=0pt,
%bar width=3pt,
bar shift=0pt
]
%
\input{hhmm/hhmm.naive-slack_times}
\input{hhmm/hhmm.histogram-slack_times}
\input{hhmm/hhmm.words-slack_times}
\input{hhmm/hhmm.time-slack_times}
%
\legend{{\footnotesize \sc Naive-Split},{\footnotesize \sc Histogram-Split},{\footnotesize \sc Words-Split},{\footnotesize \sc Times-Split}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\caption{Cumulative slack CPU time for $n$ processing cores when processing a parallel translation task split into $n$ jobs using various splitting algorithms. Slack CPU time is caused when some jobs finish before others. Zero slack time indicates conditions where all jobs complete simultaneously.}
\label{fig:slack}
\end{figure*}

@@

To observe the effects of splitting algorithms on decoding speed, we translated Urdu-English data using Moses in a parallel computing cluster, distributing work using the Sun Grid Engine. We ran two decoding setups: a standard configuration using a 5-gram language model, and a much slower syntactic LM configuration following \newcite{schwartzetal11}. Using Algorithm \ref{alg:naive}, the runtimes of the slowest of $n$ translation jobs in each configuration is illustrated in Figure \ref{fig:maxspeed} for various values of $n$. Figure \ref{fig:minmaxspeed} shows that in all cases, there is a significant difference between the fastest and slowest translation job, leading to computational slack time (Figure \ref{fig:slack}).
